---
title: "L02_ExploratoryVis"
author: "Anya Chaliotis"
date: "5/18/2019"
output:
  html_document: default
  pdf_document: default
---

## 2.  Exploratory Visualizations
Used to:  
a. Vizualize shape and spread  
b. Find patterns in your data  
c. Locate outliers  
d. summarize the main characteristics  

```{r}
library(datasets)
head(faithful)

plot(faithful, main='Old Faithful Eruptions', xlab='Eruptions length (min)', ylab='Wait time (min)', pch=20, col='blue')
```


## Histograms
Show the underlying shape of continuous data.  Allows to inspect data for outliers and overall spread. 

```{r}
hist(faithful$waiting)
```

```{r}
# Breaks (aka bins)
hist(faithful$waiting, plot=FALSE)$breaks
# how many counts in each bin
hist(faithful$waiting, plot=FALSE)$counts

#change size of bins 
hist(faithful$waiting, main='Histogram', breaks = seq(from=40,to=100,by=1))
```

## Quantile-Quantile Plots
Quantiles of the sample are plotted against quantiles of a proposed theoretical distribution.  If the points fall on a straight line, this indicates that the quantiles of the sample data are consistent with the theoretical distribution quantiles. 

```{r}
x <- rnorm(300)
qqnorm(x, pch=16, cex=.5, col='blue', main='Q-Q Plot of Normal')
qqline(x)
```

## Q-Q Plot of Faithful
Faithful dataset doesn't appear to be normally distributed.  It falls pretty far away from the normal distribution line.
```{r}
qqnorm(faithful$waiting, pch=16, cex=.5,col='blue', main='Q-Q Plot for Waiting Time')
qqline(faithful$waiting)
```

## Variability in Q-Q Plots
Smaller sample size of data that we know is normally distributed.
Still some variability in the data, eventhough we're simulating it from a normal distribution.
```{r}
x = rnorm(30); qqnorm(x); qqline(x)
```

As sample size increases, get closer to a true line but still not exact.
```{r}
x = rnorm(100); qqnorm(x); qqline(x)
```

## Barplots
```{r}
library(MASS)
data("painters")
head(painters)
```

Use the table function to create barplots
```{r}
barplot(table(painters$Composition), main='Composition Score')
barplot(table(painters$School), main='School Score')
```

